\documentclass[11pt,letterpaper]{article}
\usepackage{amsmath}
\usepackage[margin=.75in]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{appendix}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\floatstyle{ruled}
\restylefloat{table}
\restylefloat{figure}
\newcommand{\floatintro}[1]{
  \vspace*{0.1in}
  {\footnotesize
    #1
    }
  \vspace*{0.1in}
}
\newcommand{\Hline}{\noindent\rule{17cm}{0.5pt}}
\title{Homework 1 : Computational Economics \\
Search Models}
\author{Dhananjay Ghei}
\date{October 31, 2018}
\begin{document}
\maketitle
\section*{Search Models and Labor Wedge}
\begin{enumerate}
\item First show that there exists an equilibrium in the following
transformed variables: $\tilde{c}(s^t) =
c(s^t)z(s^t)^{\frac{-1}{1-\alpha}}$, $\tilde{k}(s^t) =
k(s^t)z(s^t)^{\frac{-1}{1-\alpha}}$ and $\tilde{w}(s^t) = w(s^t)
z(s^t)^{\frac{-1}{1-\alpha}}$. Then show that there exists a Markovian
solution with $k,n,s$ as state variables. Boil down the equations
below to two functional equations in two unknown functions:
$\theta(k,n,s)$ and $\tilde{c}(k,n,s)$. \\ \Hline \\
Appendix \ref{appendix:a} gives the calculations for the equations
derived below with the appropriate transformations.\\
The Euler equation is given by:
\begin{align*}
  1 = \beta \sum_{s^{t+1}|s^t} \pi(s_{t+1}|s_t)
  \frac{\tilde{c}(s^t)}{\tilde{c}(s^{t+1})}
  e^{-\frac{s_{t+1}}{1-\alpha}} \bigg(\alpha \bigg(
  \frac{\tilde{k}(s^{t+1})}{n(s^{t+1}) - \theta(s^{t+1})
  (1-n(s^{t+1}))}\bigg)^{\alpha-1} + 1 - \delta \bigg)
\end{align*}
The condition w.r.t employment is given by:
\begin{align*}
  (1-\alpha) \bigg(\frac{\tilde{k}(s^t)}{n(s^t)-\theta(s^t)(1-n(s^t))}
  \bigg)^{\alpha} = \beta \mu(\theta(s^t)) \sum_{s^{t+1}|s^t}
  \pi(s_{t+1}|s_t)\\
   \times \bigg( (1-\alpha) \bigg(\frac{\tilde{k}(s^{t+1})}{n(s^{t+1})
  - \theta (s^{t+1}) (1-n(s^{t+1}))}\bigg)^{\alpha} \bigg(1 +
  \frac{1-x}{\mu(\theta (s^{t+1}))} \bigg) - \tilde{w}(s^{t+1}) \bigg)
\end{align*}
The wage equation evaluated in history $s^{t+1}$ reduces to:
\begin{align*}
  (1-\tau) \tilde{w}(s^{t+1}) = (1-\phi) \gamma \tilde{c}(s^{t+1})
  + \phi (1-\tau)(1-\alpha) \bigg(\frac{\tilde{k}(s^{t+1})}{n(s^{t+1})
  - \theta (s^{t+1})(1-n(s^{t+1}))} \bigg)^{\alpha} (1+\theta(s^{t+1}))
\end{align*}
Eliminating wage $\tilde{w}(s^{t+1})$ between the last two equations
gives:
\begin{align*}
  (1-\alpha) \bigg(\frac{\tilde{k}(s^t)}{n(s^t) -
  \theta(s^t)(1-n(s^t))} \bigg)^{\alpha} = \beta \mu(\theta(s^t))
  \sum_{s^{t+1}|s^t} \pi(s_{t+1}|s_t)
  \frac{\tilde{c}(s^t)}{\tilde{c}(s^{t+1})}\\ \bigg(-
  \frac{(1-\phi)\gamma \tilde{c}(s^{t+1})}{1-\tau} + (1-\alpha) \bigg(
  \frac{\tilde{k}(s^{t+1})}{n(s^{t+1}) - \theta(s^{t+1})(1-n(s^{t+1}))}
  \bigg)^{\alpha} \bigg(\frac{1-x}{\mu(\theta(s^{t+1}))} + 1 -\phi -
  \phi \theta(s^{t+1}) \bigg)\bigg)
\end{align*}
The resource constraint is given by:
\begin{align*}
  \tilde{k}(s^{t+1}) e^{\frac{s_{t+1}}{1-\alpha}} =
  (\tilde{k}(s^t))^{\alpha} (n(s^t) - \theta(s^t)(1-n(s^t)))^{1-\alpha}
  + (1-\delta) \tilde{k}(s^t) -\tilde{c}(s^t)
\end{align*}
The law of motion for employment is given by:
\begin{align*}
  n(s^{t+1}) = (1-x) n(s^t) + f(\theta(s^t))(1-n(s^t))
\end{align*}
In order to show that there exists a Markovian solution, we first
conjecture that the solution is given as constants $\bar{c}, \bar{k},
\bar{w}, \bar{w}, \bar{\theta}$. Then, we
plug in these values into the 5 equations above. If there is such an
equilibrium, the equations
are given by:
\begin{align*}
  1 &= \beta \sum \pi(s_{t+1}|s_t) e^{-\frac{s_{t+1}}{1-\alpha}}
  \bigg(\alpha \bigg(\frac{\bar{k}}{\bar{n} -
  \bar{\theta}(1-\bar{n})}\bigg)^{\alpha-1} + 1 - \delta\bigg)\\
  (1-\alpha)
  \bigg(\frac{\bar{k}}{\bar{n}-\bar{\theta}(1-\bar{n})}\bigg)^{\alpha}
  &= \beta \mu(\bar{\theta}) \sum \pi(s_{t+1}|s_t)
  \times \bigg((1-\alpha) \bigg(\frac{\bar{k}}{\bar n - \bar \theta
  (1-\bar n)}\bigg)^{\alpha} \bigg(1 + \frac{1-x}{\mu(\bar \theta)}
  \bigg) - \bar{w}\bigg)\\
  \bar n &= (1-x)\bar n + f(\bar \theta) (1-\bar n)\\
  \bar k e^{\frac{s_{t+1}}{1-\alpha}} &= \bar k^{\alpha} (\bar n - \bar
  \theta (1-\bar n))^{1-\alpha} + (1-\delta) \bar k - \bar c\\
  (1-\tau) \bar w &= (1-\phi) \gamma \bar c + \phi (1-\alpha) (1-\tau)
  \bigg(\frac{\bar k}{\bar n - \bar \theta (1-\bar n)} \bigg)^{\alpha} (1+\bar \theta)
\end{align*}
Substituting out $\bar c, \bar n, \bar k$, we get an equation in
$\bar \theta$ and $\bar w$ as:
% Add equation
There is an equilibrium of this form if and only if this equation can
be satisfied in any history $s^t$. Thus, this requires that there
exists a number $\bar s$ satisfying:
% Add equation

Finally, let the functions $\Theta$ and $C$ define the
recruiter-employment ratio and equilibrium consumption relative to
trend as functions of the current state $(s,n,\tilde{k})$ so,
$\theta(s^t) = \Theta(s_t, n(s^t), \tilde{k}(s^t))$ and
$\tilde{c}(s^t) = C(s_t, n(s^t), \tilde{k}(s^t))$. Substituting in the
expressions and eliminating $\tilde{k}(s^{t+1})$ and $n(s^{t+1})$, we
get two nonlinear equations in $\Theta$ and $C$. Appendix
\ref{appendix:b} gives the calculations for this transformation. The
functional equations are given as:

\item Compute the non stochastic steady state. Log linearise around
  the non stochastic steady state to get a system of linear
  expectational difference equations and solve for the policy
  rules. For example:
  \begin{align*}
    \text{log}\theta_{t+1} = \text{log}\bar{\theta} + \theta_s
    (s_{t+1}-\bar{s}) + \theta_n (\text{log}n_t - \text{log}\bar{n}) +
    \theta_k (\text{log}k_t - \text{log}\bar{k})
  \end{align*}
\Hline \\
Appendix \ref{appendix:c} uses the equations in part 1) to derive the
equations for the non stochastic steady state. The steady state
equations are given by:

We put these equations along with the parameter values to solve for
the steady state using a non-linear solver. The steady state is given
by
\item Using the calibration in Table 3.2 to compute the IRF and the
  ergodic moments of all the relevant variables. Please check your
  results against figure 3.2 in the book. \\ \Hline \\
For most of the parameters, Shimer uses the same calibration as in
Table 3.2. However, he changes some of the parameters. He sets
$\bar{s} = 0.0012$, $\rho = 0.4$ and $\zeta = 0.00325$. 
Table \ref{tab:calibration} gives the values
of the remaining parameters.
\begin{table}[htbp!]
\label{tab:calibration}
\floatintro{These are the same parameters as in Shimer's book from
  Table 3.2}
\centering
\begin{tabular}{lrr}
\hline
Parameter & Description & Values \\
\hline
$\beta$ & Discounting & 0.996 \\
$\alpha$ & Share of capital & 0.33 \\
$\tau$ & Tax rate & 0.4 \\
$x$ & Employment exit probability & 0.034 \\
$\gamma$ & Worker's bargaining power & 0.5 \\
$\delta$ & Depreciation & 0.0028 \\
\hline
\end{tabular}
\caption{Calibration of parameters}
\end{table}
\item Derive an expression for labor wedge in this economy using the
  consumption output ratio and the hours which here is a fraction of
  HH that are employed. Obtain a log linear expansion of the wedge
  using the log linear policy rules. \\ \Hline \\
\item What is the correlation of the labor wedge with output and
  employment? You can sign it by staring at the policy rules. What is
  the ergodic std. of the labor wedge? How does the correlations and
  volatility compare to data. \\ \Hline \\
\item Now extract a TFP series such that the simulate path of $y(s^t)$
  is exactly as in the data. Start with the steady state level of
  capital and employment. Using the policy rule for output reverse
  engineer a shock such that detrended output matches exactly to the
  value in the data. Now update the state variables using your policy
  rules and keep iterating this procedure. How does the TFP series
  extracted compare to a Solow residual in the data? \\ \Hline \\ 
\end{enumerate}
\newpage
\begin{appendices}
  \section{Transforming the variables on a balanced growth
    path} \label{appendix:a}
  \section{Reducing the set of equations to a pair of nonlinear
    equations} \label{appendix:b}
  \section{Steady state calculations} \label{appendix:c}
\end{appendices}
\end{document}
